---
title: "Module 6"
author: "Erin Anderson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## CHALLENGE 1

## Load the Country-Data-2016 dataset into a data frame variable, d, and summarize the variables in that data frame. You can load the file any way you want, e.g., load from a local file, or you can access the data straight from GitHub, as in the code below.
```{r Country data 2016, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/Country-Data-2016.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
summary(d)
names(d)
```
## What are the median area and population size of all countries in the dataset?
## Create a new pop_density variable in your data frame which is population / area
## What are the 10 most dense countries? The 10 least dense?
```{r median area, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
summary(d$area)
pop_density <- d$population/d$area
d <- d[order(pop_density), ]
d[1:10, ]

```
## Extract data from the 20 largest countries into a new variable. What are the median area and population size of these countries?
## CANT DO THIS ONE
```{r 20 largest countries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Extract data from all countries beginning with the letters “A” through “F”. What are the mean area and population size of these countries?
```{r A-F, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
new <- d[grep("^[A-F]", d$country), ]
summary(new)
mean(new$population, na.rm = TRUE)
mean(new$area, na.rm = TRUE)

```

## CHALLENGE 2
```{r A-F, include=FALSE}
par(mfrow = c(2, 3))
boxplot(d$population)
boxplot(log(d$population))
boxplot(d$area)
boxplot(log(d$area))
barplot(d$population)
barplot(d$area)
```

## CHALLENGE 3
```{r A-F, include=FALSE}
par(mfrow = c(1, 2))  # gives us two panels
attach(d)
hist(log(population), freq = FALSE, col = "red", main = "Plot 1", xlab = "log(population size)",
    ylab = "density", ylim = c(0, 0.2))
hist(log(area), freq = FALSE, col = "red", main = "Plot 2", xlab = "log(area)",
    ylab = "density", ylim = c(0, 0.2))

```
# creates a histogram

## CHALLENGE 3
```{r A-F, include=FALSE}
par(mfrow = c(1, 1))  # set up one panel and redraw the log(population) histogram
hist(log(population), freq = FALSE, col = "white", main = "My Plot with Mean and Density",
+      xlab = "log(population size)", ylab = "density", ylim = c(0, 0.2))
abline(v = mean(log(population), na.rm = TRUE), col = "blue")
lines(density(log(population), na.rm = TRUE), col = "green")

```
# creates a bar chart